---
title: "R Markdown Workshop"
subtitle: "Reproducible Reports with R Markdown"
author: "Emi Tanaka"
email: "dr.emi.tanaka@gmail.com"
date: "1st Dec 2019 @ Biometrics by the Botanic Gardens | Adelaide, Australia"
length: "75 minutes"
color_theme: "yellow"
pdflink: "session4.pdf"
params:
  data: data1.csv
titleslide_bgimg: "images/bg1.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/cssgrid.css"
      - "assets/custom.css"
      - "assets/fira-code.css"
      - "assets/image_modal.css"
      - "assets/boxes.css"
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    beforeInit: "assets/custom.js"
    lib_dir: libs
    includes:
      in_header: "assets/custom.html"
      # before_body: 
      after_body: 
       - "assets/remarkjs_extended.html"
       - "assets/image_modal.js"
      # includes
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: googlecode
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---

```{r external, child="scripts/setup.Rmd", include=FALSE}
```
```{r titleslide, child="scripts/titleslide.Rmd"}
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.path = "figure/session4/"
)
```

---

class: center middle


.font_large[`r emo::ji("search")` Open and inspect the file]

.font_large[`demo-header.Rmd`]

---

# Cross Reference


::: grid 

::: item 

* When you make a header via Rmd
```markdown
# Some Header
```
 an .red[id is created automatically].
* The id is created by replacing .blue[space with `-`] and making it .blue[all lower case].
* Now you can link to this header by `[some text](#some-header)`.
* Cross references work for both pdf and html outputs.

::: 

::: item 

<center>
<iframe src="exercises/demo-header.html" width = "90%", height = "500px"></iframe>
</center>

:::

:::


---


# Direct Reference for `html`

<br>

::: paddings

* For `html` output, you can also give a link directly to the relevant section
* E.g. open `demo-header.html` in a web browser
* Append say `#chicken-data` to the url. It should look like 
.center[
`demo-header.html#chicken-data`
]
* It should have taken you to straight to the corresponding header `r emo::ji("exercise")`

:::

---


# User-defined id


<br>

::: paddings 

* You can define your own id by appending `{#your-id}`.

```markdown
# Some header {#header1}
```
* Now you can link to this header with the id `header1`.
* Note there should be no space in the id name!

::: 

---


class: center middle


.font_large[`r emo::ji("search")` Open and inspect the file]

.font_large[`demo.bib`]



---

class: font_smaller

# Bibliography

* BibTeX citation style format is used to store references in `.bib` files.
* Remember that you can get most BibTeX citation for R packages `citation` function. (Scroll below to see the BibTeX citation).

.scroll-350[
```{r}
citation("xaringan")
```
]

---

class: center middle


.font_large[`r emo::ji("search")` Open, inspect and knit the file]

.font_large[`demo-citation.Rmd`]

---

# Citations

<br>

::: paddings

* You can include BibTeX by specifying the `bib` file at YAML as:

```markdown
bibliography: bibliography.bib
```


[@bibtex-key] <i class="fas fa-arrow-right"></i> (Author et al. 2019)

or 

@bibtex-key <i class="fas fa-arrow-right"></i> Author et al. 2019

* See `demo-citation.Rmd`


:::

---

class: font_small


# Figure References

::: paddings 

* Support for figure references are included for output format type `bookdown::pdf_document2` for pdf or `bookdown::html_document2` for html.

````markdown
`r ''````{r plot1, fig.cap = "Caption"}
ggplot(cars, aes(dist, speed)) + geom_point()
```
````

* Above figure number can be referenced as .red[`\@ref(fig:plot1)`]

* The reference label has the prefix `fig:` before the chunk label.




:::

---

class: font_small


# Table References

::: paddings 

* Support for table references are also included for output format type `bookdown::pdf_document2` for pdf or `bookdown::html_document2` for html.

````markdown
`r ''````{r table1}
knitr::kable(cars, booktabs = TRUE, caption = "Caption")
```
````

* Above table number can be referenced as .red[`\@ref(tab:table1)`]

* The reference label has the prefix `tab:` before the chunk label.

:::


---

class: font_small

# Markdown for Captions

::: paddings

````markdown
`r ''````{r plot1, fig.cap = "(ref:label)"}
ggplot(cars, aes(dist, speed)) + geom_point()
```
````

* Then the caption can be entered in a separate paragraph with empty lines above and below it

```markdown

(ref:label) This is the *caption* with **markdown**.

```

* You can substitute `label` with another unique label composed of alphanumeric characters, :, -, or /
* This caption supports markdown syntax
* This is great for long captions
* It also works for tables!


:::


---


class: center middle


.font_large[`r emo::ji("wrench")` Open and work through]

.font_large[`challenge-12-references.Rmd`]

`r countdown::countdown(5, font_size = "1em")`




---

class: font_small

# Parametrized Report

````markdown
---
title: "Parameterized Report"
params:
*  species: setosa 
output: html_document
---

`r ''````{r, message = FALSE, fig.dim = c(3,2)}
library(tidyverse)
iris %>% 
*  filter(Species==params$species) %>% 
   ggplot(aes(Sepal.Length, Sepal.Width)) +
   geom_point(aes(color=Species))
```
````

::: {.output .pos top:60px; right:10px;}
<img src = "images/parameterised-report-eg1.png" width = "550px"/>

:::

---

class: font_smaller

# Knit with Parameters

::: grid
::: item 

```markdown
---
title: "Parameterized Report"
params:
  species: 
    label: "Species"
    value: setosa
    input: select
    choices: [setosa, versicolor, virginica]
  color: red
  max:
    label: "Maximum Sepal Width"
    value: 4
    input: slider
    min: 4
    max: 5
    step: 0.1
output: html_document
---
```

:::

::: item 

<img src = "images/knit-with-params.png" width = "250px" style="border: solid 3px black;"/>

````markdown
`r ''````{r, message = params$printmsg}
library(tidyverse)
iris %>% 
  filter(Species==params$species) %>% 
  filter(Sepal.Width < params$max) %>% 
  ggplot(aes(Sepal.Length, Sepal.Width)) +
  geom_point(color = params$color) + 
  labs(title = params$species)
```
````



:::

:::

---

class: font_smaller

# Shiny Report Generator
::: grid
::: item 
```markdown
---
title: "Parameterized Report"
params:
  species: 
    label: "Species"
    value: setosa
    input: select
    choices: [setosa, versicolor, virginica]
  color: red
  max:
    label: "Maximum Sepal Width"
    value: 5
    input: slider
    min: 4
    max: 5
    step: 0.05
output: html_document
---
```

:::

::: item 

<Br>

<img src = "images/shiny-params.png" width = "100%" style="border: solid 3px black;"/>

:::

:::

---

class: center middle


.font_large[`r emo::ji("wrench")` Open and work through]

.font_large[`challenge-13-params.Rmd`]

`r countdown::countdown(5, font_size = "1em")`

---

class: font_smaller

# R Markdown via Command Line

::: grid 

::: item 

`demo-render.Rmd`

````markdown
---
title: "Parameterized Report"
params:
*  species: setosa
output: html_document
---

`r ''````{r, message = FALSE, fig.dim = c(3,2)}
library(tidyverse)
iris %>% 
*  filter(Species==params$species) %>% 
   ggplot(aes(Sepal.Length, Sepal.Width)) +
   geom_point(aes(color=Species))
```
````

:::

::: item 

You can knit this file via R command by<br> using the `render` function:

```{r, eval = FALSE}
library(rmarkdown)
render("demo-render.Rmd")
```

<br>

You can overwrite the YAML values<br>  by supplying arguments to `render`:

```{r, eval = FALSE}
library(rmarkdown)
render("demo-render.Rmd", 
       output_format = "pdf_document",
       params = list(species = "virginica"))
```

:::

:::

---

class: center middle


.font_large[`r emo::ji("wrench")` Open and work through]

.font_large[`challenge-14-letters.Rmd`]

`r countdown::countdown(10, font_size = "1em")`


---

# Themes: `html_document`  

<!-- The Modal -->
<div id="myModal" class="modal">
  <div id="caption"></div>
  <span class="close" onclick="document.getElementById('myModal').style.display='none'">&times;</span>
  <iframe class="modal-content" id="img01"></iframe><br>
</div>

You can change the look of the html document by specifying themes:
::: grid 

::: item 

* `default` <img class="modalImg" alt="default"/> 
* `cerulean` <img class="modalImg" alt="cerulean"/> 
* `journal`  <img class="modalImg" alt="journal"/> 
* `flatly` <img class="modalImg" alt="flatly"/> 
* `darkly` <img class="modalImg" alt="darkly"/> 
* `readable` <img class="modalImg" alt="readable"/> 
* `spacelab` <img class="modalImg" alt="spacelab"/> 
* `united` <img class="modalImg" alt="united"/> 

:::

::: item 


* `cosmo` <img class="modalImg" alt="cosmo"/> 
* `lumen` <img class="modalImg" alt="lumen"/> 
* `paper` <img class="modalImg" alt="paper"/> 
* `sandstone` <img class="modalImg" alt="sandstone"/> 
* `simplex` <img class="modalImg" alt="simplex"/> 
* `yeti` <img class="modalImg" alt="yeti"/> 
* `NULL`  <img class="modalImg" alt="null"/> 

:::

::: item 


```markdown
output: 
  html_document:
    theme: cerulean
```

These [bootswatch](https://bootswatch.com/) themes attach the whole bootstrap library which makes your html file size larger.

:::

:::

---

# `prettydoc` 

<!-- The Modal -->
<div id="myModal2" class="modal">
  <div id="caption2"></div>
  <span class="close" onclick="document.getElementById('myModal2').style.display='none'">&times;</span>
  <iframe class="modal-content" id="img02"></iframe><br>
</div>

::: grid 

::: item 

`prettydoc` `r emo::ji('package')` is a community contributed `theme` that is light-weight:

* `cayman` <img class="modalImg2" alt="cayman"/> 
* `tactile` <img class="modalImg2" alt="tactile"/> 
* `architect` <img class="modalImg2" alt="architect"/> 
* `leonids` <img class="modalImg2" alt="leonids"/> 
* `hpstr` <img class="modalImg2" alt="hpstr"/> 



:::

::: item 



```markdown
output: 
  prettydoc::html_pretty:
    theme: cayman
```

See more about it below:

https://prettydoc.statr.me/

:::

:::





---

# `rmdformats`

<!-- The Modal -->
<div id="myModal3" class="modal">
  <div id="caption3"></div>
  <span class="close" onclick="document.getElementById('myModal3').style.display='none'">&times;</span>
  <iframe class="modal-content" id="img03"></iframe><br>
</div>


::: grid 

::: item 

`rmdformats` `r emo::ji('package')` contains four built-in `html` formats: 


* `readthedown` <img class="modalImg3" alt="readthedown"/> 
* `html_clean` <img class="modalImg3" alt="html_clean"/> 
* `html_docco` <img class="modalImg3" alt="html_docco"/> 
* `material` <img class="modalImg3" alt="material"/> 

:::

::: item 

You can use these formats by simply specifying the output in YAML as below:


```markdown
output: rmdformats::readthedown
```

See more about it below:

https://github.com/juba/rmdformats

:::

:::

---

# `rticles` - LaTeX Journal Article Templates

<!-- The Modal -->
<div id="myModal4" class="modal">
  <div id="caption4"></div>
  <span class="close" onclick="document.getElementById('myModal4').style.display='none'">&times;</span>
  <iframe class="modal-content" id="img04"></iframe><br>
</div>
::: grid
::: item 
* `acm` <img class='modalPdf' alt='acm_article'/>
* `acs` <img class='modalPdf' alt='acs_article'/>
* `aea` <img class='modalPdf' alt='aea_article'/>
* `agu` <img class='modalPdf' alt='agu_article'/>
* `amq` <img class='modalPdf' alt='amq_article'/>
* `ams` <img class='modalPdf' alt='ams_article'/>
* `asa` <img class='modalPdf' alt='asa_article'/>
* `biometrics` <img class='modalPdf' alt='biometrics_article'/>
* `copernicus` <img class='modalPdf' alt='copernicus_article'/>


:::

::: item 

* `elsevier` <img class='modalPdf' alt='elsevier_article'/>
* `frontiers` <img class='modalPdf' alt='frontiers_article'/>
* `ieee` <img class='modalPdf' alt='ieee_article'/>
* `jss` <img class='modalPdf' alt='jss_article'/>
* `mdpi` <img class='modalPdf' alt='mdpi_article'/>
* `mnras` <img class='modalPdf' alt='mnras_article'/>
* `peerj` <img class='modalPdf' alt='peerj_article'/>
* `plos` <img class='modalPdf' alt='plos_article'/>

:::

::: item 
* `pnas` <img class='modalPdf' alt='pnas_article'/>
* `rjournal` <img class='modalPdf' alt='rjournal_article'/>
* `rsos` <img class='modalPdf' alt='rsos_article'/>
* `rss` <img class='modalPdf' alt='rss_article'/>
* `sage` <img class='modalPdf' alt='sage_article'/>
* `sim` <img class='modalPdf' alt='sim_article'/>
* `springer` <img class='modalPdf' alt='springer_article'/>
* `tf` <img class='modalPdf' alt='tf_article'/>

::::

::::


::: {.bottom_abs .width100 .bg-indigo .white padding-left:10px;}
Go to RStudio > File > New File > R Markdown ... > From Template
:::


---


# External Files in Templating

::: grid

::: item 

* When using `rticles`, each journal usually require external files (e.g. `cls` or image files).
* These external components are stored within the package.
* So use `draft` instead of `render`! 

::: 

::: item bg-gray

**GUI**

* `RStudio > File > New File > R Markdown ... > From Template`

**Command line**

```{r, eval = FALSE}
rmarkdown::draft("file.Rmd", 
 template = "biometrics_article",
 package = "rticles")
```



:::

:::

---

# Making your own R Markdown template 

::: paddings

* You need to make an R package first!<br>
Go to RStudio > New Project > New Directory > R Package or `usethis::create_package()`


* When you are in your R package project,

```{r, eval = FALSE}
usethis::use_rmarkdown_template("<Name>")
```

* Modify the `skeleton/skeleton.Rmd` to how you want and add all external files to the `skeleton` folder.
* Install your package.
* `r emo::ji("tada")` And now find it at RStudio > File > New File > R Markdown > From Template.

:::

---

class: center middle


.font_large[`r emo::ji("wrench")` Create your own <br>R Markdown Template Package!]


---

```{r endslide, child="scripts/endslide.Rmd"}
```

