---
title: "R Markdown Workshop"
subtitle: "Challenge: References"
output: 
  bookdown::html_document2:
    number_sections: false
---


```{r setup, include = FALSE}
library(tidyverse)
knitr::opts_chunk$set(
  fig.height = 3.5,
  fig.width = 3.5,
  fig.align = "center", 
  echo = FALSE
)
theme_set(theme_bw(base_size = 14))
```

## Challenge

* Knit this document and see what happens to where the text contains `@\ref`
* Change this document output to `bookdown::pdf_document2`.

## Report

There is a theory that the faster a car travels, the distances taken to stop is longer!

We collected observations on `r nrow(cars)` distances taken to stop for cars at different speeds. The scatter plot of this data is shown in Figure \@ref(fig:plot1).


```{r plot1, fig.cap = "This is just a simple scatterplot"}
ggplot(cars, aes(speed, dist)) + 
  geom_point() +
  labs(x = "Speed", y = "Distance to stop") + 
  geom_smooth(method = "lm")
```

The estimate of intercept and slope of a linear model to this data can be found in Table \@ref(tab:model1).

```{r model1}
fit <- lm(dist ~ speed, cars)
knitr::kable(broom::tidy(fit), 
             booktabs = TRUE,
             caption = "The linear model fit")
```

Various goodness of fit measures for the model can be seen in Table \@ref(tab:GoF).


```{r GoF}
knitr::kable(broom::glance(fit), 
             booktabs = TRUE, 
             caption = "(ref:goodness-of-fit)")
```


(ref:goodness-of-fit) This caption is written in a *separate* paragraph. Notice **R markdown** syntax can be used! The goodness of fit measures in this table refers to the model fitted in Table \@ref(tab:model1).

(ref:resplot) The text reference also works for figures as well and can appear anywhere in the R Markdown document.

```{r resplot, fig.cap = "(ref:resplot)"}
broom::augment(fit) %>% 
  ggplot(aes(speed, .resid)) + 
  geom_point() + 
  labs(y = "Residual", x = "Speed", title = "Residual plot")

```

